---
title: "Classification Summative Coursework"
author: "bgtt65"
output:
  pdf_document: default
  html_notebook: default
---

# Data Exploration

Load the heart failure data, take a look the head and check for null values. It 
can be observed that there are some variables are in binary. We then convert 
numeric to factor.  

```{r}
heart.failure <- readr::read_csv("heart_failure.csv")
#View(heart.failure)
head(heart.failure)
is.null(heart.failure)

heart.failure$anaemia <- as.factor(heart.failure$anaemia)
heart.failure$diabetes <- as.factor(heart.failure$diabetes)
heart.failure$high_blood_pressure <- as.factor(heart.failure$high_blood_pressure)
heart.failure$sex <- as.factor(heart.failure$sex)
heart.failure$smoking <- as.factor(heart.failure$smoking)
heart.failure$fatal_mi <- as.factor(heart.failure$fatal_mi)
```

Have a look at the summary of the data. Then we briefly understand the counts of
factor variables and the distribution of numeric variables.

```{r}
skimr::skim(heart.failure)
```

As we want to predict whether the patient suffered a fatal myocardial infarction 
during the follow-up period. We plot a pie chart of the target variable.

```{r}
library("dplyr")
library("ggplot2")

fatal_mi.table <- as.data.frame(table(heart.failure$fatal_mi))
prop1 = round(fatal_mi.table[1,2] / sum(fatal_mi.table[,2]) * 100, 1)
prop2 = round(fatal_mi.table[2,2] / sum(fatal_mi.table[,2]) * 100, 1)
fatal_mi.data <- data.frame(
  HeartFailure = c("Yes", "No"),
  prop = c(prop1, prop2),
  plabel = paste(as.character(c(prop1, prop2)),"%")
)

fatal_mi.data <- fatal_mi.data %>%
  arrange(desc(HeartFailure)) %>%
  mutate(lab.ypos = cumsum(prop) - 0.5*prop)

ggplot(fatal_mi.data, aes(x = 2, y = prop, fill = HeartFailure)) +
  geom_bar(stat = "identity", color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = lab.ypos, label = plabel), color = "white") +
  scale_fill_manual(values = hcl.colors(dim(fatal_mi.data)[1], "Fall")) +
  theme_void() +
  ggtitle("Patient Status") +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
```

Show the relationship between target variable and sex.

```{r}
fatal_mi.table <- as.data.frame(table(heart.failure$fatal_mi, heart.failure$sex))
prop1 = round(fatal_mi.table[1,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop2 = round(fatal_mi.table[2,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop3 = round(fatal_mi.table[3,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop4 = round(fatal_mi.table[4,3] / sum(fatal_mi.table[,3]) * 100, 1)
fatal_mi.data <- data.frame(
  HeartFailure = c("Yes (Female)", "No (Female)", "Yes (Male)", "No (Male)"),
  prop = c(prop1, prop2, prop3, prop4),
  plabel = paste(as.character(c(prop1, prop2, prop3, prop4)),"%")
)

fatal_mi.data <- fatal_mi.data %>%
  arrange(desc(HeartFailure)) %>%
  mutate(lab.ypos = cumsum(prop) - 0.5*prop)

ggplot(fatal_mi.data, aes(x = 2, y = prop, fill = HeartFailure)) +
  geom_bar(stat = "identity", color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = lab.ypos, label = plabel), color = "white") +
  scale_fill_manual(values = hcl.colors(dim(fatal_mi.data)[1], "Fall")) +
  theme_void() +
  ggtitle("Patient Status (Sex)") +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
```

Show the relationship between target variable and anaemia.

```{r}
fatal_mi.table <- as.data.frame(table(heart.failure$fatal_mi, heart.failure$anaemia))
prop1 = round(fatal_mi.table[1,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop2 = round(fatal_mi.table[2,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop3 = round(fatal_mi.table[3,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop4 = round(fatal_mi.table[4,3] / sum(fatal_mi.table[,3]) * 100, 1)
fatal_mi.data <- data.frame(
  HeartFailure = c("Yes (Not Anaemia)", "No (Not Anaemia)", "Yes (Anaemia)", "No (Anaemia)"),
  prop = c(prop1, prop2, prop3, prop4),
  plabel = paste(as.character(c(prop1, prop2, prop3, prop4)),"%")
)

fatal_mi.data <- fatal_mi.data %>%
  arrange(desc(HeartFailure)) %>%
  mutate(lab.ypos = cumsum(prop) - 0.5*prop)

ggplot(fatal_mi.data, aes(x = 2, y = prop, fill = HeartFailure)) +
  geom_bar(stat = "identity", color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = lab.ypos, label = plabel), color = "white") +
  scale_fill_manual(values = hcl.colors(dim(fatal_mi.data)[1], "Fall")) +
  theme_void() +
  ggtitle("Patient Status (Anaemia)") +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
```

Show the relationship between target variable and diabetes.

```{r}
fatal_mi.table <- as.data.frame(table(heart.failure$fatal_mi, heart.failure$diabetes))
prop1 = round(fatal_mi.table[1,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop2 = round(fatal_mi.table[2,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop3 = round(fatal_mi.table[3,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop4 = round(fatal_mi.table[4,3] / sum(fatal_mi.table[,3]) * 100, 1)
fatal_mi.data <- data.frame(
  HeartFailure = c("Yes (Not Diabetes)", "No (Not Diabetes)", "Yes (Diabetes)", "No (Diabetes)"),
  prop = c(prop1, prop2, prop3, prop4),
  plabel = paste(as.character(c(prop1, prop2, prop3, prop4)),"%")
)

fatal_mi.data <- fatal_mi.data %>%
  arrange(desc(HeartFailure)) %>%
  mutate(lab.ypos = cumsum(prop) - 0.5*prop)

ggplot(fatal_mi.data, aes(x = 2, y = prop, fill = HeartFailure)) +
  geom_bar(stat = "identity", color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = lab.ypos, label = plabel), color = "white") +
  scale_fill_manual(values = hcl.colors(dim(fatal_mi.data)[1], "Fall")) +
  theme_void() +
  ggtitle("Patient Status (Diabetes)") +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
```

Show the relationship between target variable and high blood pressure.

```{r}
fatal_mi.table <- as.data.frame(table(heart.failure$fatal_mi, heart.failure$high_blood_pressure))
prop1 = round(fatal_mi.table[1,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop2 = round(fatal_mi.table[2,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop3 = round(fatal_mi.table[3,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop4 = round(fatal_mi.table[4,3] / sum(fatal_mi.table[,3]) * 100, 1)
fatal_mi.data <- data.frame(
  HeartFailure = c("Yes (Not HBP)", "No (Not HBP)", "Yes (HBP)", "No (HBP)"),
  prop = c(prop1, prop2, prop3, prop4),
  plabel = paste(as.character(c(prop1, prop2, prop3, prop4)),"%")
)

fatal_mi.data <- fatal_mi.data %>%
  arrange(desc(HeartFailure)) %>%
  mutate(lab.ypos = cumsum(prop) - 0.5*prop)

ggplot(fatal_mi.data, aes(x = 2, y = prop, fill = HeartFailure)) +
  geom_bar(stat = "identity", color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = lab.ypos, label = plabel), color = "white") +
  scale_fill_manual(values = hcl.colors(dim(fatal_mi.data)[1], "Fall")) +
  theme_void() +
  ggtitle("Patient Status (High Blood Pressure)") +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
```

Show the relationship between target variable and smoking.

```{r}
fatal_mi.table <- as.data.frame(table(heart.failure$fatal_mi, heart.failure$smoking))
prop1 = round(fatal_mi.table[1,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop2 = round(fatal_mi.table[2,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop3 = round(fatal_mi.table[3,3] / sum(fatal_mi.table[,3]) * 100, 1)
prop4 = round(fatal_mi.table[4,3] / sum(fatal_mi.table[,3]) * 100, 1)
fatal_mi.data <- data.frame(
  HeartFailure = c("Yes (Not Smoking)", "No (Not Smoking)", "Yes (Smoking)", "No (Smoking)"),
  prop = c(prop1, prop2, prop3, prop4),
  plabel = paste(as.character(c(prop1, prop2, prop3, prop4)),"%")
)

fatal_mi.data <- fatal_mi.data %>%
  arrange(desc(HeartFailure)) %>%
  mutate(lab.ypos = cumsum(prop) - 0.5*prop)

ggplot(fatal_mi.data, aes(x = 2, y = prop, fill = HeartFailure)) +
  geom_bar(stat = "identity", color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = lab.ypos, label = plabel), color = "white") +
  scale_fill_manual(values = hcl.colors(dim(fatal_mi.data)[1], "Fall")) +
  theme_void() +
  ggtitle("Patient Status (Smoking)") +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
```

Show the bar charts for factor variables.

```{r}
library(ggplot2)
DataExplorer::plot_bar(heart.failure, ncol = 2, ggtheme = theme_linedraw())
DataExplorer::plot_bar(heart.failure, by = "fatal_mi", ncol = 2, ggtheme = theme_linedraw())
```

Show the histograms for numeric variables.

```{r}
DataExplorer::plot_histogram(heart.failure, ncol = 3, ggtheme = theme_linedraw())
```

Show the Q-Q plots for numeric variables.

```{r}
DataExplorer::plot_qq(heart.failure, ncol = 3, ggtheme = theme_linedraw())
DataExplorer::plot_qq(heart.failure, by = "fatal_mi", ncol = 3, ggtheme = theme_linedraw())
```

Show the box plots for numeric variables.

```{r}
DataExplorer::plot_boxplot(heart.failure, by = "fatal_mi", ncol = 4, ggtheme = theme_linedraw())
```

Show the correlation for numeric variables.

```{r}
DataExplorer::plot_correlation(heart.failure, type = "continuous")
```

## Model Fitting

Firstly, load `tidyverse`, MLR3 and related tools.

```{r}
library("tidyverse")
# data structure used by MLR3
library("data.table")
library("mlr3verse")
```

Make a heart failure task.

```{r}
set.seed(500) 
heartfailure_task <- TaskClassif$new(id = "HeartFailure",
                               backend = heart.failure,
                               target = "fatal_mi",
                               positive = "1")
```

Set several options for good error estimation.

```{r}
holdout <- rsmp("holdout", ratio = 0.7)
holdout$instantiate(heartfailure_task)

cv <- rsmp("cv", folds = 5)
cv$instantiate(heartfailure_task)

bootstrap <- rsmp("bootstrap")
bootstrap$instantiate(heartfailure_task)
```

Set several algorithms for model fitting comparison.

```{r}
lrn_baseline <- lrn("classif.featureless", predict_type = "prob")
lrn_log_reg <- lrn("classif.log_reg", predict_type = "prob")
lrn_cart <- lrn("classif.rpart", predict_type = "prob")
lrn_xgboost <- lrn("classif.xgboost", predict_type = "prob")
pl_xgb <- po("encode") %>>%
  po(lrn_xgboost)
```

Set the measurements for loss function.

```{r}
measures <- list(msr("classif.ce"),
                 msr("classif.acc"),
                 msr("classif.auc"),
                 msr("classif.fpr"),
                 msr("classif.fnr"))
```

Set the benchmarks and show the results of measurements.

```{r}
library("precrec")

# test/train split
res <- benchmark(data.table(
  task       = list(heartfailure_task),
  learner    = list(lrn_baseline,
                    lrn_log_reg,
                    lrn_cart,
                    pl_xgb),
  resampling = list(holdout)
), store_models = TRUE)

res$aggregate(measures)
autoplot(res)
autoplot(res$clone(deep = TRUE), type = "roc")

# cross validation
res <- benchmark(data.table(
  task       = list(heartfailure_task),
  learner    = list(lrn_baseline,
                    lrn_log_reg,
                    lrn_cart,
                    pl_xgb),
  resampling = list(cv)
), store_models = TRUE)

res$aggregate(measures)
autoplot(res)
autoplot(res$clone(deep = TRUE), type = "roc")

# bootstrap
res <- benchmark(data.table(
  task       = list(heartfailure_task),
  learner    = list(lrn_baseline,
                    lrn_log_reg,
                    lrn_cart),
  resampling = list(bootstrap)
), store_models = TRUE)

res$aggregate(measures)
autoplot(res)
autoplot(res$clone(deep = TRUE), type = "roc")
```

```{r}
library("precrec")

# cross validation
res <- benchmark(data.table(
  task       = list(heartfailure_task),
  learner    = list(lrn_log_reg,
                    lrn_cart),
  resampling = list(cv)
), store_models = TRUE)

autoplot(res$clone(deep = TRUE), type = "roc")
```

## Model Improvements

Use feature selection wrappers

```{r}
library("mlr3fselect")

set.seed(500)
instance <- fselect(
  method = "random_search",
  task =  heartfailure_task,
  learner = lrn("classif.log_reg"),
  resampling = rsmp("cv", folds = 5),
  measure = msr("classif.ce"),
  term_evals = 10,
  batch_size = 5
)
instance$result
```

Build a new dataset without some irrelevant variables that we observed from data
exploration and .

```{r}
re_heart.failure <- heart.failure %>% 
  select(-creatinine_phosphokinase, -platelets)
```

Make a new heart failure task (without some irrelevant variables).

```{r}
set.seed(500) 
re_heartfailure_task <- TaskClassif$new(id = "ReHeartFailure",
                                  backend = re_heart.failure,
                                  target = "fatal_mi",
                                  positive = "1")
```

Set several options for good error estimation.

```{r}
re_holdout <- rsmp("holdout", ratio = 0.7)
re_holdout$instantiate(re_heartfailure_task)

re_cv <- rsmp("cv", folds = 5)
re_cv$instantiate(re_heartfailure_task)

re_bootstrap <- rsmp("bootstrap")
re_bootstrap$instantiate(re_heartfailure_task)
```

Set the benchmarks and show the results of measurements.

```{r}
res <- benchmark(data.table(
  task       = list(re_heartfailure_task),
  learner    = list(lrn_log_reg,
                    lrn_cart),
  resampling = list(re_cv)
), store_models = TRUE)

res$aggregate(measures)
autoplot(res)
autoplot(res$clone(deep = TRUE), type = "roc")

res <- benchmark(data.table(
  task       = list(re_heartfailure_task),
  learner    = list(lrn_log_reg,
                    lrn_cart),
  resampling = list(re_bootstrap)
), store_models = TRUE)

res$aggregate(measures)
autoplot(res)
autoplot(res$clone(deep = TRUE), type = "roc")
```

Examine in depth the results in each fold of cross validation.

```{r}
plot(res$resample_result(2)$learners[[1]]$model, compress = TRUE, margin = 0.1)
text(res$resample_result(2)$learners[[1]]$model, use.n = TRUE, cex = 0.8)

plot(res$resample_result(2)$learners[[2]]$model, compress = TRUE, margin = 0.1)
text(res$resample_result(2)$learners[[2]]$model, use.n = TRUE, cex = 0.8)

plot(res$resample_result(2)$learners[[3]]$model, compress = TRUE, margin = 0.1)
text(res$resample_result(2)$learners[[3]]$model, use.n = TRUE, cex = 0.8)

plot(res$resample_result(2)$learners[[4]]$model, compress = TRUE, margin = 0.1)
text(res$resample_result(2)$learners[[4]]$model, use.n = TRUE, cex = 0.8)

plot(res$resample_result(2)$learners[[5]]$model, compress = TRUE, margin = 0.1)
text(res$resample_result(2)$learners[[5]]$model, use.n = TRUE, cex = 0.8)
```

Make a selection for the cost penalty to prune the trees.

```{r}
lrn_cart_cv <- lrn("classif.rpart", predict_type = "prob", xval = 10)

res_cart_cv <- resample(re_heartfailure_task, lrn_cart_cv, re_cv, store_models = TRUE)
rpart::plotcp(res_cart_cv$learners[[1]]$model)
rpart::plotcp(res_cart_cv$learners[[2]]$model)
rpart::plotcp(res_cart_cv$learners[[3]]$model)
rpart::plotcp(res_cart_cv$learners[[4]]$model)
rpart::plotcp(res_cart_cv$learners[[5]]$model)
```

Set the benchmarks and show the results of measurements.

```{r}
lrn_cart_cp <- lrn("classif.rpart", predict_type = "prob", cp = 0.15)

res <- benchmark(data.table(
  task       = list(re_heartfailure_task),
  learner    = list(lrn_log_reg,
                    lrn_cart_cp),
  resampling = list(re_cv)
), store_models = TRUE)

res$aggregate(measures)
autoplot(res)
autoplot(res$clone(deep = TRUE), type = "roc")
autoplot(res, type = "prc")
```

## Performance Report

Create some performance plots.

```{r}
library("precrec")

res <- benchmark(data.table(
  task       = list(re_heartfailure_task),
  learner    = list(lrn_log_reg),
  resampling = list(re_cv)
), store_models = TRUE)

autoplot(res)
autoplot(res$clone(deep = TRUE), type = "roc")
autoplot(res, type = "prc")

library(cvms)
library(tibble)

measures_cm <- list(msr("classif.tp"),
                    msr("classif.fp"), 
                    msr("classif.fn"),
                    msr("classif.tn"))

cfm_data <- res$aggregate(measures_cm)

tp <- as.integer(cfm_data[1,7] / sum(cfm_data[1,7:10]) * dim(re_heart.failure)[1])
fp <- as.integer(cfm_data[1,8] / sum(cfm_data[1,7:10]) * dim(re_heart.failure)[1])
fn <- as.integer(cfm_data[1,9] / sum(cfm_data[1,7:10]) * dim(re_heart.failure)[1])
tn <- as.integer(cfm_data[1,10] / sum(cfm_data[1,7:10]) * dim(re_heart.failure)[1])

cfm_table <- matrix(c(tn, fp, fn, tp), ncol=2, byrow=TRUE)
colnames(cfm_table) <- c("0","1")
rownames(cfm_table) <- c("0","1")
names(dimnames(cfm_table)) <- c("target", "prediction")
cfm_table <- as.table(cfm_table)
cfm_table
cfm <- as_tibble(cfm_table)
plot_confusion_matrix(cfm, 
                      target_col = "target", 
                      prediction_col = "prediction",
                      counts_col = "n")
```