---
title: "GEOL50130 - Mini Project"
author: "bgtt65"
output:
  pdf_document: default
  html_notebook: default
---

# Part 1: Data Exploration

Load the Durham temperature data first and take a look at the head and check for 
null values. The year, month and day variables are in numeric, then we convert 
numeric to factor.  

```{r}
library(tidyverse)

# Load data and omit null values
durhamTemp <- readr::read_csv("durhamtemp_1901_2019.csv")
head(durhamTemp)
durhamTemp <- na.omit(durhamTemp)

# Convert numeric to factor
durhamTemp$Year <- as.factor(durhamTemp$Year)
durhamTemp$Month <- as.factor(durhamTemp$Month)
durhamTemp$Day <- as.factor(durhamTemp$Day)
```

Show the summary of the Durham temperature data. Now we briefly understand the 
distribution of temperature and PPT and the counts of date variables.

```{r}
# Review the data
skimr::skim(durhamTemp)
```

```{r}
library(ggplot2)

# Plot the histograms for temperature and PPT
DataExplorer::plot_histogram(durhamTemp$`Av temp`, ncol = 2, ggtheme = theme_linedraw())
```

```{r}
# Plot the Q-Q plots for temperature and PPT
DataExplorer::plot_qq(durhamTemp$`Av temp`, ncol = 2, ggtheme = theme_linedraw())
```

```{r}
library(oce)
library(dplyr)

# Filter the values of 29th Feb in every four years
durhamTemp.fil <- filter(durhamTemp, Month!='2' | Day!='29')

# Set the total number of days and years
days = 1:365
years = 1901:2019

# Make a matrix, then plot a heatmap with contour
durhamTemp.mat = matrix(durhamTemp.fil$`Av temp`, nrow = length(days)) 
imagep(x = years, y = days, z = durhamTemp.mat %>% t(), filledContour = F, 
       col = oceColors9A(120), ylim = c(365,0),
       ylab = "Day", zlab = expression(Temperature~(degree *C)))
contour(x = years, y = days, z = durhamTemp.mat %>% t(), add = TRUE, 
        col = 1, nlevels = 3)
```

```{r}
# Make the daily time series object and plot
daily.ts = ts(data = durhamTemp$`Av temp`, start = c(1901,1,1), frequency = 365.24)
plot(daily.ts, xlab = "Time", las = 1, ylab = expression(Temperature~(degree*C)),
     main = "Durham Daily Tempterature Time Series")
```

Make the monthly average dataset and plot the monthly time series.

```{r}
library(lubridate)
library(forecast)

# Summarise the data to monthly average
durhamTemp.monthly <- durhamTemp%>%
  group_by(Year,Month)%>%
  summarise(totalRain = sum(PPT.), meanAvtemp = mean(`Av temp`),
            meanTmax = mean(Tmax), meanTmin = mean(Tmin))
durhamTemp.monthly$Date <- paste("15/",durhamTemp.monthly$Month,"/",durhamTemp.monthly$Year)
durhamTemp.monthly$Date <- dmy(durhamTemp.monthly$Date)

# Make the monthly time series object and plot
monthly.ts = ts(data = durhamTemp.monthly$meanAvtemp, start = c(1901,1), frequency = 12)
plot(monthly.ts, xlab = "Time", las = 1, ylab = expression(Temperature~(degree*C)),
     main = "Durham Monthly Tempterature Time Series")

# Make a season plot
ggseasonplot(monthly.ts, year.labels=TRUE, year.labels.left=TRUE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Durham Monthly Tempterature")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

# Make a subseries plot
ggsubseriesplot(monthly.ts)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Durham Monthly Tempterature")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
```

# Part 2: Monthly Temperature Forecast 

```{r}
# Make training (80%) and test (20%) sets
monthly.ts.train <- window(monthly.ts, start = c(1901,1), end = c(1995,12)) 
monthly.ts.test <- window(monthly.ts, start = c(1996,1), end = c(2019,12)) 

# Make cross-validation set
monthly.ts.CV <- window(monthly.ts, start = c(2015,1), end = c(2019,12)) 
```

## 2.1 Basic Methods

**Mean Method**

```{r}
## Strategy 1: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.mean <- meanf(monthly.ts.train, h = 12 * 24)
accuracyTT.mean <- accuracy(forecastTT.mean, monthly.ts.test)
rmseTT.mean <- accuracy(forecastTT.mean, monthly.ts.test)[2,2]
checkresiduals(forecastTT.mean)

## Strategy 2: Time series cross-validation
# forecast the value of test set temperature and compute the accuracy(RMSE)
errorCV.mean <- tsCV(monthly.ts.CV, forecastfunction = meanf, h = 6)
mseCV.mean <-colMeans(errorCV.mean^2, na.rm = T)
ggplot(data.frame(h = 1:6, MSE = mseCV.mean), aes(x = h, y = MSE))+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of Mean Method Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
rmseCV.mean <- sqrt(mean(mseCV.mean))
```

**Naïve Method**

```{r}
## Strategy 1: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.naive <- naive(monthly.ts.train, h = 12 * 24)
accuracyTT.naive <- accuracy(forecastTT.naive, monthly.ts.test)
rmseTT.naive <- accuracy(forecastTT.naive, monthly.ts.test)[2,2]
checkresiduals(forecastTT.naive)

## Strategy 2: Time series cross-validation
# forecast the value of test set temperature and compute the accuracy(RMSE)
errorCV.naive <- tsCV(monthly.ts.CV, forecastfunction = rwf, h = 6)
mseCV.naive <-colMeans(errorCV.naive^2, na.rm = T)
ggplot(data.frame(h = 1:6, MSE = mseCV.naive), aes(x = h, y = MSE))+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of Naïve Method Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
rmseCV.naive <- sqrt(mean(mseCV.naive))
```

**Seasonal Naïve Method**

```{r}
## Strategy 1: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.snaive <- snaive(monthly.ts.train, h = 12 * 24)
accuracyTT.snaive <- accuracy(forecastTT.snaive, monthly.ts.test)
rmseTT.snaive <- accuracy(forecastTT.snaive, monthly.ts.test)[2,2]
checkresiduals(forecastTT.snaive)

## Strategy 2: Time series cross-validation
# forecast the value of test set temperature and compute the accuracy(RMSE)
errorCV.snaive <- tsCV(monthly.ts.CV, forecastfunction = snaive, h = 6)
mseCV.snaive <-colMeans(errorCV.snaive^2, na.rm = T)
ggplot(data.frame(h = 1:6, MSE = mseCV.snaive), aes(x = h, y = MSE))+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of Snaïve Method Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
rmseCV.snaive <- sqrt(mean(mseCV.snaive))
```

**Draft Method**

```{r}
## Strategy 1: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.draft <- rwf(monthly.ts.train, h = 12 * 24, drift = TRUE, na.rm=TRUE)
accuracyTT.draft <- accuracy(forecastTT.draft, monthly.ts.test)
rmseTT.draft <- accuracy(forecastTT.draft, monthly.ts.test)[2,2]
checkresiduals(forecastTT.draft)

## Strategy 2: Time series cross-validation
# forecast the value of test set temperature and compute the accuracy(RMSE)
errorCV.draft <- tsCV(monthly.ts.CV, forecastfunction = rwf, drift = TRUE, h = 6)
mseCV.draft <-colMeans(errorCV.draft^2, na.rm = T)
ggplot(data.frame(h = 1:6, MSE = mseCV.draft), aes(x = h, y = MSE))+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of Draft Method Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
rmseCV.draft <- sqrt(mean(mseCV.draft))
```

**Bsics Methods Comparison**

After the comparison with four basic methods, the seasonal naïve shows a good
performance. This method is chosen as the best basic model to set a baseline
for other following models.

```{r}
## Strategy 1: Training and test sets
# show accuracy results for these basic methods
accuracyTT.basic <- as.data.frame(rbind(accuracyTT.mean, accuracyTT.naive,
                                        accuracyTT.snaive, accuracyTT.draft))
rownames(accuracyTT.basic) <- c("Mean Training", "Mean Test", "Naïve Training", "Naïve Test",
                                "Snaïve Training", "Snaïve Test", "Draft Training", "Draft Test")
accuracyTT.basic 

# plot forecast results for these basic methods
autoplot(monthly.ts)+
  autolayer(forecastTT.snaive, series = "Saïve", PI = FALSE)+
  autolayer(forecastTT.mean, series = "Mean", PI = FALSE)+
  autolayer(forecastTT.naive, series = "Naïve", PI = FALSE)+
  autolayer(forecastTT.draft, series = "Draft", PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Basic Methods Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  guides(colour = guide_legend(title = "Method"))+
  scale_color_manual(values = c("#ef1828", "#3f60aa", "#f46f20", "#9ec417"))

## Strategy 2: Time series cross-validation
# plot accuracy results for these basic methods
mseCV.basic <- data.frame(h = 1:6, mseCV.mean = mseCV.mean, mseCV.naive = mseCV.naive,
                          mseCV.snaive = mseCV.snaive, mseCV.draft = mseCV.draft)
ggplot(mseCV.basic)+
  geom_point(aes(x = h, y = mseCV.mean, colour = "Mean"), size=2)+
  geom_point(aes(x = h, y = mseCV.naive, colour = "Naïve"), size=2)+
  geom_point(aes(x = h, y = mseCV.snaive, colour = "Snaïve"), size=2)+
  geom_point(aes(x = h, y = mseCV.draft, colour = "Draft"), size=2)+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of Basic Methods Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  guides(colour = guide_legend(title = "Method"))+
  scale_color_manual(values = c("#ef1828", "#3f60aa", "#f46f20", "#9ec417"))

# show accuracy results for these basic methods
accuracyCV.basic <- as.data.frame(rbind(rmseCV.mean, rmseCV.naive,
                                        rmseCV.snaive, rmseCV.draft))
colnames(accuracyCV.basic) <- c("RMSE")
rownames(accuracyCV.basic) <- c("Mean", "Naïve", "Snaïve", "Draft")
accuracyCV.basic 
```

## 2.2 Linear Regression Model

```{r}
library('fpp')

## Strategy 1: Training and test sets
# forecast the value of test set temperature and compute the accuracy
fitTT.tslm <- tslm(monthly.ts.train ~ trend + season)
forecastTT.tslm <- forecast(fitTT.tslm, h = 12 * 24)
accuracyTT.tslm <- accuracy(forecastTT.tslm, monthly.ts.test)
rmseTT.tslm <- accuracy(forecastTT.tslm, monthly.ts.test)[2,2]
checkresiduals(forecastTT.tslm)

# plot forecast results for linear regression model
autoplot(monthly.ts)+
  autolayer(forecastTT.tslm, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Linear Regression Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))

## Strategy 2: Time series cross-validation
# forecast the value of test set temperature and compute the accuracy(RMSE)
tslm_function <- function(data, h) {
  fitmodel <- tslm(data ~ trend + season)
  forecast(fitmodel, h = h)
}
errorCV.tslm <- tsCV(monthly.ts.CV, tslm_function, h = 6)
mseCV.tslm <-colMeans(errorCV.tslm^2, na.rm = T)
ggplot(data.frame(h = 1:6, MSE = mseCV.tslm), aes(x = h, y = MSE))+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of Linear Regression Model Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
rmseCV.tslm <- sqrt(mean(mseCV.tslm))
```

## 2.3 Exponential Smoothing

**ETS (Error, Trend, Seasonal) Model**

```{r}
## Strategy 1: Training and test sets
# forecast the value of test set temperature and compute the accuracy
fitTT.ets <- ets(monthly.ts.train)
forecastTT.ets <- forecast(fitTT.ets, h = 12 * 24)
accuracyTT.ets <- accuracy(forecastTT.ets, monthly.ts.test)
rmseTT.ets <- accuracy(forecastTT.ets, monthly.ts.test)[2,2]
checkresiduals(forecastTT.ets)

# plot forecast results for ETS model
autoplot(monthly.ts)+
  autolayer(forecastTT.ets, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("ETS(A,N,A) Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))

## Strategy 2: Time series cross-validation
# forecast the value of test set temperature and compute the accuracy(RMSE)
ets_function <- function(data, h) {
  fitmodel <- ets(data)
  forecast(fitmodel, h = h)
}
errorCV.ets <- tsCV(monthly.ts.CV, ets_function, h = 6)
mseCV.ets <-colMeans(errorCV.ets^2, na.rm = T)
ggplot(data.frame(h = 1:6, MSE = mseCV.ets), aes(x = h, y = MSE))+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of ETS Model Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
rmseCV.ets <- sqrt(mean(mseCV.ets))
```

## 2.4 ARIMA Model

```{r}
## Strategy 1: Training and test sets
# forecast the value of test set temperature and compute the accuracy
fitTT.arima <- auto.arima(monthly.ts.train)
forecastTT.arima <- forecast(fitTT.arima, h = 12 * 24)
accuracyTT.arima <- accuracy(forecastTT.arima, monthly.ts.test)
rmseTT.arima <- accuracy(forecastTT.arima, monthly.ts.test)[2,2]
checkresiduals(forecastTT.arima)

# plot forecast results for ARIMA model
autoplot(monthly.ts)+
  autolayer(forecastTT.arima, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("ARIMA(1,0,0)(2,1,0) Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))

## Strategy 2: Time series cross-validation
# forecast the value of test set temperature and compute the accuracy(RMSE)
arima_function <- function(data, h) {
  fitmodel <- auto.arima(data)
  forecast(fitmodel, h = h)
}
errorCV.arima <- tsCV(monthly.ts.CV, arima_function, h = 6)
mseCV.arima <-colMeans(errorCV.arima^2, na.rm = T)
ggplot(data.frame(h = 1:6, MSE = mseCV.arima), aes(x = h, y = MSE))+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of ARIMA Model Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
rmseCV.arima <- sqrt(mean(mseCV.arima))
```

## 2.5 Decomposition

**STL + ETS Model**

```{r}
library(fpp2)

## Strategy 1: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.stle <- stlf(monthly.ts.train, h = 12 * 24, method = "ets")
accuracyTT.stle <- accuracy(forecastTT.stle, monthly.ts.test)
rmseTT.stle <- accuracy(forecastTT.stle, monthly.ts.test)[2,2]
checkresiduals(forecastTT.stle)

# plot forecast results for STL model
autoplot(monthly.ts)+
  autolayer(forecastTT.stle, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("STL + ETS(A,N,N) Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))

## Strategy 2: Time series cross-validation
# forecast the value of test set temperature and compute the accuracy(RMSE)
stle_function <- function(data, h) {
  stlf(data, h = h, method = "ets")
}
errorCV.stle <- tsCV(monthly.ts.CV, stle_function, h = 6)
mseCV.stle <-colMeans(errorCV.stle^2, na.rm = T)
ggplot(data.frame(h = 1:6, MSE = mseCV.stle), aes(x = h, y = MSE))+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of STL + ETS Model Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
rmseCV.stle <- sqrt(mean(mseCV.stle))
```

**STL + ARIMA Model**

```{r}
library(fpp2)

## Strategy 1: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.stla <- stlf(monthly.ts.train, h = 12 * 24, method = "arima")
accuracyTT.stla <- accuracy(forecastTT.stla, monthly.ts.test)
rmseTT.stla <- accuracy(forecastTT.stla, monthly.ts.test)[2,2]
checkresiduals(forecastTT.stla)

# plot forecast results for STL model
autoplot(monthly.ts)+
  autolayer(forecastTT.stla, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("STL + ARIMA(1,1,2) Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))

## Strategy 2: Time series cross-validation
# forecast the value of test set temperature and compute the accuracy(RMSE)
stla_function <- function(data, h) {
  stlf(data, h = h, method = "arima")
}
errorCV.stla <- tsCV(monthly.ts.CV, stla_function, h = 6)
mseCV.stla <-colMeans(errorCV.stla^2, na.rm = T)
ggplot(data.frame(h = 1:6, MSE = mseCV.stla), aes(x = h, y = MSE))+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of STL + ARIMA Model Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
rmseCV.stla <- sqrt(mean(mseCV.stla))
```

## 2.6 Neural Network Model

```{r}
## Strategy 1: Training and test sets
# forecast the value of test set temperature and compute the accuracy
set.seed(2022)
fitTT.nnetar <- nnetar(monthly.ts.train)
forecastTT.nnetar <- forecast(fitTT.nnetar,h = 12 * 24)
accuracyTT.nnetar <- accuracy(forecastTT.nnetar, monthly.ts.test)
rmseTT.nnetar <- accuracy(forecastTT.nnetar, monthly.ts.test)[2,2]
checkresiduals(forecastTT.nnetar)

# plot forecast results for NN model
autoplot(monthly.ts)+
  autolayer(forecastTT.nnetar, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Neural Network Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))

## Strategy 2: Time series cross-validation
# forecast the value of test set temperature and compute the accuracy(RMSE)
nnetar_function <- function(data, h) {
  set.seed(2022)
  fitmodel <- nnetar(data)
  forecast(fitmodel, h = h)
}
errorCV.nnetar <- tsCV(monthly.ts.CV, nnetar_function, h = 6)
mseCV.nnetar <-colMeans(errorCV.nnetar^2, na.rm = T)
ggplot(data.frame(h = 1:6, MSE = mseCV.nnetar), aes(x = h, y = MSE))+
  geom_point()+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of Neural Network Model Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
rmseCV.nnetar <- sqrt(mean(mseCV.nnetar))
```

## 2.7 Model Comparison

```{r}
## Strategy 1: Training and test sets
# show accuracy results for all comparison models
accuracyTT.all <- as.data.frame(rbind(accuracyTT.snaive, accuracyTT.tslm, 
                                      accuracyTT.ets, accuracyTT.arima,
                                      accuracyTT.stle, accuracyTT.stla,
                                      accuracyTT.nnetar))
rownames(accuracyTT.all) <- c("Snaïve Training", "Snaïve Test", "LR Training", "LR Test",
                              "ETS Training", "ETS Test", "ARIMA Training", "ARIMA Test",
                              "STL+ETS Training", "STL+ETS Test", "STL+ARIMA Training", "STL+ARIMA Test", 
                              "NN Training", "NN Test")
accuracyTT.all 

# plot forecast results for all comparison models
plotTT.snaive <- autoplot(forecastTT.snaive) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.tslm <- autoplot(forecastTT.tslm) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.ets <- autoplot(forecastTT.ets) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.stle <- autoplot(forecastTT.stle) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.stla <- autoplot(forecastTT.stla) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.arima <- autoplot(forecastTT.arima) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.nnetar <- autoplot(forecastTT.nnetar) + ylab(expression(Temperature~(degree*C))) + theme_bw()
gridExtra::grid.arrange(plotTT.snaive, plotTT.tslm, plotTT.ets, nrow = 3)
gridExtra::grid.arrange(plotTT.arima, plotTT.stle, plotTT.stla, nrow = 3)
gridExtra::grid.arrange(plotTT.nnetar, nrow = 3)

## Strategy 2: Time series cross-validation
# plot accuracy results for all comparison models
mseCV.all <- data.frame(h = 1:6, mseCV.snaive = mseCV.snaive, mseCV.tslm = mseCV.tslm,
                        mseCV.ets = mseCV.ets, mseCV.arima = mseCV.arima,
                        mseCV.stle = mseCV.stle, mseCV.stla = mseCV.stla,
                        mseCV.nnetar)
ggplot(mseCV.basic)+
  geom_point(aes(x = h, y = mseCV.snaive, colour = "Snaïve"), size = 2)+
  geom_point(aes(x = h, y = mseCV.tslm, colour = "LR"), size = 2)+
  geom_point(aes(x = h, y = mseCV.ets, colour = "ETS"), size = 2)+
  geom_point(aes(x = h, y = mseCV.arima, colour = "ARIMA"), size = 2)+
  geom_point(aes(x = h, y = mseCV.stle, colour = "STL+ETS"), size = 2)+
  geom_point(aes(x = h, y = mseCV.stla, colour = "STL+ARIMA"), size = 2)+
  geom_point(aes(x = h, y = mseCV.nnetar, colour = "NN"), size = 2)+
  scale_x_continuous(breaks = seq(1, 6))+
  xlab("Number of periods")+
  ylab("MSE")+
  ggtitle("MSE of Basic Methods Using Cross-vaildation")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  guides(colour = guide_legend(title = "Method"))+
  scale_color_manual(values = c("#ef1828", "#3f60aa", "#f46f20", "#9ec417", "gray", "gold", "purple"))

# show accuracy results for all comparison models
accuracyCV.all <- as.data.frame(rbind(rmseCV.snaive, rmseCV.tslm, rmseCV.ets, rmseCV.arima,
                                      rmseCV.stle, rmseCV.stla, rmseCV.nnetar))
colnames(accuracyCV.all) <- c("RMSE")
rownames(accuracyCV.all) <- c("Snaïve", "LR", "ETS", "ARIMA", "STL+ETS", "STL+ARIMA", "NN")
accuracyCV.all
```

## 2.8 Prediction

The STL + ETS model is the best model after comparison, then use this model to 
make a 2020 monthly temperature prediction.

```{r}
# forecast the value of 2020 temperature
forecastTT.stl.2020 <- stlf(monthly.ts.train, h = 12 * 25, method = "ets")
prediction.stl.test <- window(forecastTT.stl.2020$mean, start = c(1996,1), end = c(2019,12)) 
prediction.stl.2020 <- window(forecastTT.stl.2020$mean, start = c(2020,1), end = c(2020,12)) 

# plot forecast results for the best model
autoplot(monthly.ts)+
  autolayer(prediction.stl.test, series = "Test Set")+
  autolayer(prediction.stl.2020, series = "2020")+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Temperature Forecast Using STL + ETS Model")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
  guides(colour = guide_legend(title = "Forecast"))
  
# plot 2020 forecast for the best model
autoplot(prediction.stl.2020)+
  scale_x_yearmon(format = "%b", n = 15)+
  geom_point(size = 3)+
  geom_text(aes(label = round(prediction.stl.2020, 1)),
            vjust = 0.5, hjust = -0.3)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("2020 Temperature Forecast Using STL + ETS Model")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```

# Part 3: Daily Temperature Forecast 

```{r}
# Make training (80%) and test (20%) sets
daily.ts.sets <- window(daily.ts, start = c(2010,1), end = c(2019,366)) 
daily.ts.train <- window(daily.ts, start = c(2010,1), end = c(2017,366)) 
daily.ts.test <- window(daily.ts, start = c(2018,1), end = c(2019,366)) 
```

## 3.1 Basic Methods

**Mean Method**

```{r}
## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.mean <- meanf(daily.ts.train, h = 2 * 365.24)
accuracyTT.mean <- accuracy(forecastTT.mean, daily.ts.test)
rmseTT.mean <- accuracy(forecastTT.mean, daily.ts.test)[2,2]
checkresiduals(forecastTT.mean)
```

**Naive Method**

```{r}
## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.naive <- naive(daily.ts.train, h = 2 * 365.24)
accuracyTT.naive <- accuracy(forecastTT.naive, daily.ts.test)
rmseTT.naive <- accuracy(forecastTT.naive, daily.ts.test)[2,2]
checkresiduals(forecastTT.naive)
```

**Seasonal Naive Method**

```{r}
## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.snaive <- snaive(daily.ts.train, h = 2 * 365.24)
accuracyTT.snaive <- accuracy(forecastTT.snaive, daily.ts.test)
rmseTT.snaive <- accuracy(forecastTT.snaive, daily.ts.test)[2,2]
checkresiduals(forecastTT.snaive)
```

**Draft Method**

```{r}
## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.draft <- rwf(daily.ts.train, h = 2 * 365.24, drift = TRUE, na.rm=TRUE)
accuracyTT.draft <- accuracy(forecastTT.draft, daily.ts.test)
rmseTT.draft <- accuracy(forecastTT.draft, daily.ts.test)[2,2]
checkresiduals(forecastTT.draft)
```

**Bsics Methods Comparison**

Same as the monthly temperature forecast result, the seasonal naïve method is the 
best basic model which is set as a baseline.

```{r}
## Strategy: Training and test sets
# show accuracy results for these basic methods
accuracyTT.basic <- as.data.frame(rbind(accuracyTT.mean, accuracyTT.naive,
                                        accuracyTT.snaive, accuracyTT.draft))
rownames(accuracyTT.basic) <- c("Mean Training", "Mean Test", "Naïve Training", "Naïve Test",
                                "Snaïve Training", "Snaïve Test", "Draft Training", "Draft Test")
accuracyTT.basic 

# plot forecast results for these basic methods
autoplot(daily.ts.sets)+
  autolayer(forecastTT.snaive, series = "Saïve", PI = FALSE)+
  autolayer(forecastTT.mean, series = "Mean", PI = FALSE)+
  autolayer(forecastTT.naive, series = "Naïve", PI = FALSE)+
  autolayer(forecastTT.draft, series = "Draft", PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Basic Methods Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  guides(colour = guide_legend(title = "Method"))+
  scale_color_manual(values = c("#ef1828", "#3f60aa", "#f46f20", "#9ec417"))
```

## 3.2 Linear Regression Model (Has problem)

```{r}
library('fpp')

## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
fitTT.tslm <- tslm(daily.ts.train ~ trend + season)
forecastTT.tslm <- forecast(fitTT.tslm, h = 2 * 365.24)
accuracyTT.tslm <- accuracy(forecastTT.tslm, daily.ts.test)
rmseTT.tslm <- accuracy(forecastTT.tslm, daily.ts.test)[2,2]
checkresiduals(forecastTT.tslm)

# plot forecast results for linear regression model
autoplot(daily.ts.sets)+
  autolayer(forecastTT.tslm, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Linear Regression Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```

## 3.3 Exponential Smoothing (Has problem)

**ETS (Error, Trend, Seasonal) Model**

```{r}
## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
fitTT.ets <- ets(daily.ts.train)
forecastTT.ets <- forecast(fitTT.ets, h = 2 * 365.24)
accuracyTT.ets <- accuracy(forecastTT.ets, daily.ts.test)
rmseTT.ets <- accuracy(forecastTT.ets, daily.ts.test)[2,2]
checkresiduals(forecastTT.ets)

# plot forecast results for ETS model
autoplot(daily.ts.sets)+
  autolayer(forecastTT.ets, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("ETS(A,N,N) Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```

## 3.4 ARIMA Model (Really slow)

```{r}
## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
fitTT.arima <- auto.arima(daily.ts.train)
forecastTT.arima <- forecast(fitTT.arima, h = 2 * 365.24)
accuracyTT.arima <- accuracy(forecastTT.arima, daily.ts.test)
rmseTT.arima <- accuracy(forecastTT.arima, daily.ts.test)[2,2]
checkresiduals(forecastTT.arima)

# plot forecast results for ARIMA model
autoplot(daily.ts.sets)+
  autolayer(forecastTT.arima, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("ARIMA(3,0,1)(0,1,0) Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```

## 3.5 Decomposition

**STL + ETS Model**

```{r}
library(fpp2)

## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.stle <- stlf(daily.ts.train, h = 2 * 365.24, method = "ets")
accuracyTT.stle <- accuracy(forecastTT.stle, daily.ts.test)
rmseTT.stle <- accuracy(forecastTT.stle, daily.ts.test)[2,2]
checkresiduals(forecastTT.stle)

# plot forecast results for STL model
autoplot(daily.ts.sets)+
  autolayer(forecastTT.stle, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("STL + ETS(A,N,N) Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```

**STL + ARIMA Model**

```{r}
library(fpp2)

## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.stla <- stlf(daily.ts.train, h = 2 * 365.24, method = "arima")
accuracyTT.stla <- accuracy(forecastTT.stla, daily.ts.test)
rmseTT.stla <- accuracy(forecastTT.stla, daily.ts.test)[2,2]
checkresiduals(forecastTT.stla)

# plot forecast results for STL model
autoplot(daily.ts.sets)+
  autolayer(forecastTT.stla, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("STL + ARIMA(1,1,3) Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```

## 3.6 Neural Network Model (Has problem)

```{r}
## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
set.seed(2022)
fitTT.nnetar <- nnetar(daily.ts.train)
forecastTT.nnetar <- forecast(fitTT.nnetar,h = 2 * 365.24)
accuracyTT.nnetar <- accuracy(forecastTT.nnetar, daily.ts.test)
rmseTT.nnetar <- accuracy(forecastTT.nnetar, daily.ts.test)[2,2]
checkresiduals(forecastTT.nnetar)

# plot forecast results for NN model
autoplot(daily.ts.sets)+
  autolayer(forecastTT.nnetar, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Neural Network Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```

## 3.7 Model Comparison

```{r}
## Strategy: Training and test sets
# show accuracy results for all comparison models
accuracyTT.all <- as.data.frame(rbind(accuracyTT.snaive, accuracyTT.tslm, 
                                      accuracyTT.ets, accuracyTT.arima,
                                      accuracyTT.stle, accuracyTT.stla,
                                      accuracyTT.nnetar))
rownames(accuracyTT.all) <- c("Snaïve Training", "Snaïve Test", "LR Training", "LR Test",
                              "ETS Training", "ETS Test", "ARIMA Training", "ARIMA Test",
                              "STL+ETS Training", "STL+ETS Test", "STL+ARIMA Training", "STL+ARIMA Test", 
                              "NN Training", "NN Test")
accuracyTT.all 

# plot forecast results for all comparison models
plotTT.snaive <- autoplot(forecastTT.snaive) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.tslm <- autoplot(forecastTT.tslm) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.ets <- autoplot(forecastTT.ets) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.stle <- autoplot(forecastTT.stle) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.stla <- autoplot(forecastTT.stla) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.arima <- autoplot(forecastTT.arima) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.nnetar <- autoplot(forecastTT.nnetar) + ylab(expression(Temperature~(degree*C))) + theme_bw()
gridExtra::grid.arrange(plotTT.snaive, plotTT.arima, plotTT.stle, nrow = 3)
gridExtra::grid.arrange(plotTT.stla, nrow = 3)
```

## 3.8 Model Refitting

```{r}
# Make training (80%) and test (20%) sets
daily.ts.train <- window(daily.ts, start = c(1901,1), end = c(1995,366)) 
daily.ts.test <- window(daily.ts, start = c(1996,1), end = c(2019,366)) 
```

**Seasonal Naive Method**

```{r}
## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.snaive <- snaive(daily.ts.train, h = 24 * 365.24)
accuracyTT.snaive <- accuracy(forecastTT.snaive, daily.ts.test)
rmseTT.snaive <- accuracy(forecastTT.snaive, daily.ts.test)[2,2]
checkresiduals(forecastTT.snaive)

# plot forecast results for Seasonal naive
autoplot(daily.ts)+
  autolayer(forecastTT.snaive, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Seasonal Naive Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```

**STL + ARIMA Model**

```{r}
library(fpp2)

## Strategy: Training and test sets
# forecast the value of test set temperature and compute the accuracy
forecastTT.stl <- stlf(daily.ts.train, h = 24 * 365.24, method = "arima")
accuracyTT.stl <- accuracy(forecastTT.stl, daily.ts.test)
rmseTT.stl <- accuracy(forecastTT.stl, daily.ts.test)[2,2]
checkresiduals(forecastTT.stl)

# plot forecast results for STL model
autoplot(daily.ts)+
  autolayer(forecastTT.stl, PI = FALSE)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("STL + ETS(A,N,N) Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```

**Comparison**

```{r}
## Strategy: Training and test sets
# show accuracy results for all comparison models
accuracyTT.all <- as.data.frame(rbind(accuracyTT.snaive, accuracyTT.stl))
rownames(accuracyTT.all) <- c("Snaïve Training", "Snaïve Test", "STL+ARIMA Training", "STL+ARIMA Test")
accuracyTT.all 

# plot forecast results for all comparison models
plotTT.snaive <- autoplot(forecastTT.snaive) + ylab(expression(Temperature~(degree*C))) + theme_bw()
plotTT.stl <- autoplot(forecastTT.stl) + ylab(expression(Temperature~(degree*C))) + theme_bw()
gridExtra::grid.arrange(plotTT.snaive, plotTT.stl, nrow = 3)
```

## 3.9 Prediction

After the accuracy comparison, the STL + ARIMA model is chosen to make a 2020 daily 
temperature prediction.

```{r}
# forecast the value of 2020 temperature
forecastTT.stl.2020 <- stlf(daily.ts.train, h = 25 * 366, method = "arima")
prediction.stl.test <- window(forecastTT.stl.2020$mean, start = c(1996,1), end = c(2019,366)) 
prediction.stl.2020 <- window(forecastTT.stl.2020$mean, start = c(2020,1), end = c(2020,367)) 
prediction.stl.2020

# plot forecast results for the chosen model
autoplot(daily.ts)+
  autolayer(prediction.stl.test, series = "Test Set")+
  autolayer(prediction.stl.2020, series = "2020")+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Temperature Forecast Using STL + ARIMA Model")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
  guides(colour = guide_legend(title = "Forecast"))
  
# plot 2020 forecast for the chosen model
autoplot(prediction.stl.2020)+
  scale_x_yearmon(format = "%b", n = 16)+
  geom_point(size = 1)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("2020 Temperature Forecast Using STL + ARIMA Model")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```

# 3.10 Searching New method

**Prophet Model**

```{r}
library(prophet)

## Strategy: Training and test sets
# forecast the value of test set temperature
date <- as.character(dmy(durhamTemp$Date))
daily.ts.train.pro <- data.frame(cbind(date, durhamTemp$`Av temp`))[1:34698,]
names(daily.ts.train.pro) <- c('ds', 'y')
fitTT.pro <- prophet(daily.ts.train.pro)
future <- make_future_dataframe(fitTT.pro, periods = 8766)
forecastTT.pro <- predict(fitTT.pro, future)

# compute the accuracy
prediction.train <- forecastTT.pro['yhat'][1:34699,]
real.train <- durhamTemp$`Av temp`[1:34699]
rmseTT.pro.train <- sqrt(mean((real.train - prediction.train)^2))
rmseTT.pro.train
prediction.test <- forecastTT.pro['yhat'][34699:dim(forecastTT.pro)[1],]
real.test <- durhamTemp$`Av temp`[34699:dim(durhamTemp)[1]]
rmseTT.pro.test <- sqrt(mean((real.test - prediction.test)^2))
rmseTT.pro.test

# plot forecast results for Prophet model
plot(fitTT.pro, forecastTT.pro)+
  xlab("Time")+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("Prophet Model Forecast Using Training/Test Split")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
# forecast the value of 2020 temperature
date <- as.character(dmy(durhamTemp$Date))
future <- make_future_dataframe(fitTT.pro, periods = 9132)
forecastTT.pro.2020 <- predict(fitTT.pro, future)
prediction.pro.2020 <- forecastTT.pro.2020['yhat'][(dim(forecastTT.pro.2020)[1]-365):dim(forecastTT.pro.2020)[1],]
prediction.pro.2020 = ts(data = prediction.pro.2020, start = c(2020,1,1), frequency = 366)

# plot 2020 forecast for the chosen model
autoplot(prediction.pro.2020)+
  scale_x_yearmon(format = "%b", n = 16)+
  geom_point(size = 1)+
  ylab(expression(Temperature~(degree*C)))+
  ggtitle("2020 Temperature Forecast Using Prophet Model")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))
```
